---
title: "Assigment 3 Task 3"
author: "Carly Norris"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= FALSE, warning = FALSE)
library(janitor)
library(tidyverse)
library(here)
library(kableExtra)
library(lubridate)
```


## Read in urchins data and clean names

```{r}
urchins_sb <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Data visualization

- create a histogram to compare size distributions between purple and red urchins
- create a boxplot with common name on the x-axis and size on the y-axis

```{r}
ggplot(data= urchins_sb, aes(x= size))+
  geom_histogram()+
  facet_wrap(~common_name)+
  labs(x= "Urchin Diameter (cm)", y= "Number of Urchins")
```

```{r}
ggplot(data=urchins_sb, aes(x= common_name, y= size))+
  geom_boxplot()+
  labs(x= "Urchin Type (Common Name)", y= "Diameter (cm)")
```

## Urchin Descriptive statistics

- Create a summary table that contains mean, median, standard deviation, and sample size
- Use the kableExtra package to modify the table visualization

```{r}
urchins_summary <- urchins_sb %>% 
  group_by(common_name) %>% 
  summarize(mean_size= mean(size),
            median_size= median(size),
            sd_size= sd(size),
            sample_size= n())

kbl(urchins_summary) %>% 
  kable_material("hover")
```

How do the size (distribution, central tendency, spread, etc.) compare between the two species of urchins, ignoring site and treatment variables. 

- Purple Urchins are typically smaller than red urchins, but red urchins have a larger distribution. Red urchin mean diameter is 34% larger than purple urchin mean diameter. The sample size of purple urchins is larger than red urchin. 

## Purple Urchin Counts by Site and Year

```{r}
purple_urchins <- urchins_sb %>% 
  filter(common_name== "Purple Urchin") %>% 
  mutate(Date = mdy(date)) %>% 
  mutate(year= year(Date)) %>% 
  mutate(site_full= case_when(
    site== "NAPL" ~ "Naples Reef",
    site== "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
```

### Create a summary table containing total number of purple urchings observed by site and year

```{r}
count_purple_urchins <- purple_urchins %>% 
  count(year, site_full)

kbl(count_purple_urchins) %>% 
  kable_minimal()
```

### Exploratory graph of of purple urchin counts by site and year

```{r}
ggplot(data= count_purple_urchins, aes(x= year, y= n))+
  geom_line(aes(color=site_full))
```











